Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-03 causal-bandit/icml-2015/general.tex

Let $a \in \calA$ be an intervention and define random variable
\eq{
Z_a(X) = R_a(X) Y \ind{R_a(X) \leq B_a}\,,
}
where $B_a \geq 0$ is some constant to be chosen subsequently and
\eq{
R_a(X) = \frac{\P{\parents{Y}(X)|a}}{\Q{\parents{Y}(X)}}\,.
}
Define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}

@@

Algorithm \ref{alg:general} depends on a choice of sampling distribution $\operatorname{Q}$ that is determined by $\eta$. In light of Theorem \ref{thm:general}
a natural choice of $\eta$ is the minimiser of $m(\eta)$.
\eq{
\eta^* 
&= \argmin_\eta m(\eta) \\
&= \argmin_\eta \max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] \\
&= \argmin_\eta \underbrace{\max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\sum_{b \in \calA} \eta_b \Pn{b}{\parents{Y}(X)}}\right]}_{m(\eta)}\,.
}
Since the mixture of convex functions is convex and the maximum of a set of convex functions is convex, we see that $m(\eta)$ is convex (in $\eta$).
Therefore the minimisation problem may be tackled using standard techniques from convex optimisation. In the experimental section we present some special cases,
but for now we give two simple results. The first shows that $|\calA|$ serves as an upper bound on $m(\eta^*)$.

\begin{lemma}
$m(\eta^*) \leq |\calA|$.
\end{lemma}

@@

\todo{give reference}
The second observation is that in the parallel bandit setting the $m(\eta)$ given in this section coincides with the $m$ in Equation blah.
Recall in that setting that 
\eq{
\actions = \set{do()} \cup \set{do(X_i = j) : 1 \leq i \leq N \text{ and } j \in \set{0,1}}\,.
}
For $a = do()$ we let $\eta_a = 1/2$. Assume without loss of generality that $\P{X_i = 1} \geq \P{X_i = 0}$ for all $i$ and
$\P{X_i = 0}$ is monotone increasing in $i$. Then let $m = \max\set{i : \P{X_i = 0} \leq 1 / i}$, which is the same as in \S3. 
For $i \leq m$ and $a = do(X_i = 0)$ let $\eta_a = 1/(2m)$. Then it is easy to check that $m(\eta) \leq 2m$.

>>>>

Let $a \in \calA$ be an intervention and define random variable
\eq{
Z_a(X) = R_a(X) Y \ind{R_a(X) \leq B_a}\,,
}
where $B_a \geq 0$ is some constant to be chosen subsequently and
\eq{
R_a(X) = \frac{\P{\parents{Y}(X)|a}}{\Q{\parents{Y}(X)}}\,.
}
Abusing notation we define $m(\eta)$ by
\eq{
m(\eta) = \max_{a \in \calA} \EE_{\Ps_a}\left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right]\,.
}
We will shortly see that $m(\eta)$ is a measure of the difficulty of the problem that approximately cooincides 
with the version given in the previous section, which justifies the name overloading.

@@

Algorithm \ref{alg:general} depends on a choice of sampling distribution $\operatorname{Q}$ that is determined by $\eta$. In light of Theorem \ref{thm:general}
a natural choice of $\eta$ is the minimiser of $m(\eta)$.
\eq{
\eta^* 
&= \argmin_\eta m(\eta) \\
&= \argmin_\eta \max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\Q{\parents{Y}(X)}}\right] \\
&= \argmin_\eta \underbrace{\max_{a \in \calA} \EE_{\Ps_a} \left[\frac{\Pn{a}{\parents{Y}(X)}}{\sum_{b \in \calA} \eta_b \Pn{b}{\parents{Y}(X)}}\right]}_{m(\eta)}\,.
}
Since the mixture of convex functions is convex and the maximum of a set of convex functions is convex, we see that $m(\eta)$ is convex (in $\eta$).
Therefore the minimisation problem may be tackled using standard techniques from convex optimisation. 
An interpretation of $m(\eta^*)$ is the minimum achievable worst-case variance of the importance weighted estimator.
In the experimental section we present some special cases,
but for now we give two simple results. The first shows that $|\calA|$ serves as an upper bound on $m(\eta^*)$.

\begin{proposition}
$m(\eta^*) \leq |\calA|$.
\end{proposition}

@@

The second observation is that in the parallel bandit setting the $m(\eta)$ given in this section approximately coincides with the $m(\vec{q})$ in Eq.\ \ref{eq:m-simple}.
Recall in that setting that 
\eq{
\actions = \set{do()} \cup \set{do(X_i = j) : 1 \leq i \leq N \text{ and } j \in \set{0,1}}\,.
}
For $a = do()$ choose $\eta_a = 1/2$. 
Let $m = m(\vec{q})$ and for $a = do(X_i = j)$ let $\eta_a = \ind{\P{X_i = j} \leq 1/m} / (2m)$. 
Then it is easy to check that $m(\eta) \leq 2m$.
